---
title: "MNLM: Multiple Neutral Landscape Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MNLM-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

The MNLM package can be downloaded and installed from Github.

First install the required dependencies, using the remotes package.

install.packages("remotes")

 remotes::install_github("cran/RandomFieldsUtils")
 
 remotes::install_github("cran/RandomFields")
 
 remotes::install_github("ropensci/NLMR")

Then install MNLM from the Github repository.

 remotes::install_github("ANON_URL/MNLM")

# Running MNLM

First load the MNLM package.
```{r}
library(MNLM)
```

Correlated NLM raster layers can be generated using any of the five NLM algorithms implemented in the MNLM package using just a single line of code. For example, to generate a pair of 50 x 50 rasters with a correlation of r = 0.65 using the fractal Brownian motion algorithm, we can use the mnlm_fbm() function.
```{r, warning = FALSE}
fbm <- mnlm_fbm(nlayers = 2, r = 0.65, ncol = 50, nrow = 50, fract_dim = 1.5)
terra::plot(fbm)
```

To generate more than two NLMs in a set, simply set the 'nlayers' argument to the desired number of NLMs and include a vector with length nlayers-1 for the correlations in the 'r' argument. For example, to generate a set of three 60 x 40 rasters with correlations of r = 0.65 and 0.35 using the Gaussian random fields algorithm, we can use the mnlm_gaussianfield() function.
```{r}
gaussian <- mnlm_gaussianfield(nlayers = 3, r = c(0.65, 0.25), ncol = 60, nrow = 40)
terra::plot(gaussian)
```

Generating a set of four layers is just as simple. For example, we can use the Perlin noise algorithm implemented in the mnlm_perlin() function. Here, the 'frequency' and 'octaves' arguments specify the paramters for how the Perlin noise is generated. These can be tweaked to achieve different outcomes.
```{r}
perlin <- mnlm_perlin(nlayers = 4, r = c(0.8, 0.6, 0.4), nrow = 50, ncol = 50,
                      frequency = 0.05, octaves = 4)
terra::plot(perlin)
```

MNLM also includes the midpoint displacement algorith, which produces NLMs with structures similar to the fractal Brownian motion method. For comparison, we can generate NLMs with the mnlm_mpd() function.
```{r, warning = FALSE}
mpd <- mnlm_mpd(nlayers = 2, r = 0.65, ncol = 50, nrow = 50)
terra::plot(mpd)
```

The final NLM algorithm implemented in MNLM is the planar gradient, which is called with the mnlm_planargradient() function.
```{r}
planar <- mnlm_planargradient(nlayers = 2, r = 0.7, ncol = 50, nrow = 50)
terra::plot(planar)
```

# Custom parameterizations of individual layers

The NLM algorithms implemented in MNLM also allow for the specification of different parameters for individual layers. 

For example, we can create a set of three fractal Brownian motion layers that have fractal dimensions of 1, 0.5, and 2 by specifying a vector of these values for the 'fract_dim' argument in the mnlm_fbm() function.
```{r, warning = FALSE}
fbm <- mnlm_fbm(nlayers = 3, r = 0.65, ncol = 50, nrow = 50, fract_dim = c(1, 0.5, 2))
terra::plot(fbm)
```

And we can change both the octaves and frequency in the perlin noise algorithm to see how they change the structure of the resulting NLM layers by passing vectors to the 'frequency' and 'octaves' arguments in the mnlm_perlin() function.
```{r}
perlin <- mnlm_perlin(nlayers = 4, r = 0.5, nrow = 50, ncol = 50,
                      frequency = c(0.1, 0.3, 0.1, 0.3), octaves = c(2, 2, 6, 6))
terra::plot(perlin, main = c("freq = 0.1, octaves = 2", "freq = 0.3, octaves = 2", 
                             "freq = 0.1, octaves = 6", "freq = 0.3, octaves = 6"))
```

